Synthesizing Symmetric Lenses

### 2018-07-10 papers/SymmetricLenses/main.tex
a pass through section 5

% begin synthesis
\section{Synthesis}
\label{sec:synthesis}
Algorithm~\ref{alg:synth-sym-lens} presents our synthesis algorithm at a high
level of abstraction. The input regular expressions are first converted into
stochastic regular expressions with \ToStochastic.  
%Once that conversion has
%taken place, the $\RXSearch$ algorithm executes. This algorithm itself is
%defined via five functions: (1) $\PCF{\PQ.Create}$, (2) $\PCF{\PQ.Pop}$, (3)
%$\PCF{Expand}$, (4) $\PCF{\PQ.Push}$, and (5) $\PCF{\Continue}$. First,
This pair of SREs are used to initialize a priority queue ($pq$).
Next, $\PCF{SynthSymLens}$
enters a loop that searches for likely lenses. The loop terminates when the
algorithm believes it is unlikely to find a better lens than the best one it has
found so far (a termination condition defined by $\PCF{\Continue}$). Within each
iteration of the loop, it:
\begin{itemize}
\item pops the next class ($S$, $T$) of lenses to
search of the priority queue ($\PCF{\PQ.Pop}$),
\item executes $\GreedySynth$ to find a best lens in that class if
one exists ($\LC$),
using the examples $exs$ to filter out potential lenses that do not satisfy
the specification,
\item replaces $best$ with $\LC$, if $\LC$ has lower cost according
to our information theoretic metric, and
\item adds all of the SREs derived from rewriting $S$ and $T$ ($\Expand(S,T)$)
  to the priority queue with $\PCF{\PQ.Push}$
\end{itemize}
When the loop terminates, the search returns the globally best lens found
($best$).  Each element of this algorithm will be explained in
further depth in the following subsections.

@@

\subsection{Search Through Types}
The first synthesizer is one that searches through regular expression pairs.
This synthesizer, comprising \Expand, \PQ, and \RXSearch, proposes regular
expressions, semantically but not syntactically equivalent to the originals. In
particular, this synthesizer performs the following rewrites:
\begin{center}
  \begin{tabular}{rcl}
    $\PRegexStar{\Regex}{\Probability}$
    & \Rewrite
    & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability}\\

@@

\begin{trivlist}
  \centering
\item 
  % \begin{center}
  \begin{tabular}{@{\ }v@{\ }q}
    \LanguageOf{\PRegexStar{\DNFRegex{}}{\Probability}} \ =\  &
                                            \{\String_1\Concat\ldots\Concat\String_n
                                            \SuchThat \forall i \String_i\in\LanguageOf{\DNFRegex}\}\\
    \LanguageOf{\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n}}\ =\  & 
    % \\
    % \multicolumn{2}{L}{\ \ \ \ 
    \{\String_0\Concat\StringAlt_1\cdots\StringAlt_n\Concat\String_n \SuchThat \StringAlt_i\in\LanguageOf{\Atom_i}\}
    % }
    \\
    \LanguageOf{\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}}\ =\  &
    % \\
    % \multicolumn{2}{L}{\ \ \ \ 
    \{\String \SuchThat \String \in \LanguageOf{\Sequence_i} \text{\ and $i\in\RangeIncInc{1}{n}$}\}
    % }
  \end{tabular}
  % \end{center}
\end{trivlist}

@@

We have a means of turning stochastic regular expressions into stochastic DNF
regular expressions. The conversion algorithm itself, written
$\ToDNFRegexOf{\Regex}$, is defined below.
\[
  \begin{array}{rcl}
    \ToDNFRegexOf{\String} & = & \DNFOf{(\SequenceOf{\String},1)}\\
    \ToDNFRegexOf{\emptyset} & = & \DNFOf{}\\
    \ToDNFRegexOf{(\PRegexStar{\Regex}{\Probability})} & = & \AtomToDNFOf{\PRegexStar{(\ToDNFRegexOf{\Regex})}{\Probability}}\\
    \ToDNFRegexOf{(\RegexConcat{\Regex_1}{\Regex_2})} & = & \ToDNFRegexOf{\Regex_1} \ConcatDNF \ToDNFRegexOf{\Regex_2}\\
    \ToDNFRegexOf{(\PRegexOr{\Regex_1}{\Regex_2}{\Probability})} & = & \ToDNFRegexOf{\Regex_1} \OrDNF_{\Probability} \ToDNFRegexOf{\Regex_2}\\
  \end{array}
\]
We have proven this conversion respects languages and probability
distributions.

@@

\paragraph*{\ToStochastic} With stochastic DNF regular expressions and
\ToDNFRegex defined, it is easier to explain the function that converts regular
expressions into stochastic regular expressions, \ToStochastic. If users
specified stochastic regular expressions themselves (perhaps inferred from
example data), this step could be skipped,
though our implementation does not (yet) permit users to provide such
stochastic regular expressions. If $\Regex$ is a stochastic DNF regular
expression generated by \ToStochastic, then $\ToDNFRegexOf{\Regex} =
\DNFOf{(\Sequence_1,\frac{1}{n}) \DNFSep \ldots \DNFSep
  (\Sequence_n,\frac{1}{n})}$ for some sequences $\Sequence_1\ldots\Sequence_n$,
and every stochastic atom generated by \ToStochastic is
$\PRegexStar{\DNFRegex}{.8}$. In particular, $\ToDNFRegex$ generates regular
expressions whose DNF form gives equal probability to all sequence subcomponents
of the DNF regular expression., and gives iterations a .8 chance to continue
iterating. In our experience generating random strings from regular expressions,
these probabilities provide good distributions of strings -- stars are iterated
4 times on average, and no individual choice in a series disjunctions is chosen
disproportionately often. As we have found these distributions generate
reasonable strings, we believe they describe the shape of data well.

@@

As an example, consider the symmetric DNF lens:
$(\SDNFLensOf{(1,1,\SSQLens_1)\DNFLSep (1,2,\SSQLens_2)}
,\ListOf{2},\ListOf{1;2})$, synchronizing data between $\DNFOf{\Sequence_1}$ and
$\DNFOf{\SequenceAlt_1 \DNFSep \SequenceAlt_2}$. This lens merges both
$\SequenceAlt_1$ and $\SequenceAlt_2$ into $\Sequence_1$, as they both are
involved in a lens with $\Sequence_1$. When performing a \CreateR, $\SSQLens_2$
is used. When performing a \CreateL on a string in $\Sequence_1$, $\SSQLens_1$
is used, and when performing a \CreateL on a string in $\Sequence_2$,
$\SSQLens_2$ is used. More generally, a symmetric DNF Lens \SDNFLens{} consists
of three components. First is the mapping component. The mapping component is a
list of triples $(i,j,\SSQLens)$. In this, $i$ and $j$ refer to sequence $i$ and
sequence $j$ in the left and right DNF regular expressions, respectively. The
third component of the triple, \SSQLens{}, is a sequence lens that maps between
the two sequences specified by $i$ and $j$. The second component is a list of
integers $c$. The integer, $c_i$ at position $i$ states that when performing a
$\CreateR$ on a string matching sequence $\Sequence_i$, it will use the sequence
lens that maps it to sequence $\SequenceAlt_{c_i}$ on the right. The third
component is a list of integers $d$, and follows the same pattern as the second
component, but maps from right to left.

@@

\subsection{\GreedySynth}
\label{subsec:greedy-synth}
The greedy algorithm is actually two greedy algorithms, one that greedily finds
symmetric DNF lenses (\GreedySynth) and one that greedily finds symmetric
sequence lenses (\GreedySeqSynth). We describe this synthesis algorithm in the
context of trying to find a lens between \lstinline{"" | name.name*} and
\lstinline{"" | name}.

\paragraph*{Symmetric DNF Lenses} Algorithm~\ref{alg:greedysynth} presents
\GreedySynth, a means to synthesize symmetric DNF lenses. The inputs to
\GreedySynth are a pair of stochastic DNF regular expressions and a suite of
input-output examples.

First, \PCF{CannotMap} checks if there could possibly be a lens satisfying the
examples. If there is not, \GreedySynth returns \None early. After this,
\GreedySynth finds the best lenses between each sequence pair of the left and
right DNF regular expressions, given the examples that match them. A priority
queue containing these sequence lenses is then initialized with
$\PCF{PQ.Create}$. The symmetric lens is built up iteratively from these
sequence lenses, the state of this partial lens is tracked in the lens builder,
$lb$.

\GreedySynth then loops until there are no more sequence lenses in the priority
queue. Within this loop, sequence lenses are popped. If this sequence lens is
``useful'', then that sequence lens is included in the final DNF lens. If the
source or target of a popped sequence lens is not the source or target of
another sequence lens, then this sequence lens is considered useful. If the
examples require two sequences have a lens between them, then any sequence lens
between them is considered useful. We update the priorities of our sequence
lenses as the algorithm proceeds: if two sequence lenses have the same source,
the second one to be popped gets a higher cost than it originally had, to
address the information that needs to be stored for including that source of
non-bijectivity.

In our example, \GreedySynth may first pop the sequence pair \lstinline{""} and
\lstinline{""}, as there is a bijective sequence lens between them. As neither
\lstinline{""} is involved in a sequence lens, this lens is considered useful.
After this, the sequence lens between \lstinline{name.name*} and
\lstinline{name} would be popped: while that lens is not bijective: it is still
better than the alternatives. As all sequences are now involved in sequence lenses,
and there are no examples to make other lenses useful, no further sequence
lenses would be added to the lens builder.

@@

\paragraph*{Symmetric Sequence lenses} Algorithm~\ref{alg:greedyseqsynth}
presents \GreedySeqSynth, a means to synthesizes symmetric sequence lenses.

This algorithm uses \PCF{AtomSynth}, a simple function that synthesizes atom
lenses by iterating the DNF lens between its subcomponents. The inputs to
\GreedySeqSynth are a pair of stochastic atoms and a suite of input-output
examples. First, \PCF{CannotMap} checks if there is a possibility of a lens
between the given sequences regular expressions that is consistent with the
examples. If there is not, \GreedySeqSynth returns \None early. After this,
\GreedySeqSynth finds the best lenses between each atom pair of the left and
right sequences. A priority queue containing these lenses is then initialized
with \PCF{PQ.Create}. The symmetric sequence lens is built up iteratively from
these atom lenses, the state of the partial lens is tracked in the sequence lens
builder, $slb$.

\GreedySeqSynth then loops until there are no more atom lenses in the priority
queue. Within this loop, atom lenses are popped. The atom lens is then added if
it is useful. An atom lens is useful if adding it will lower the cost of the
generated sequence lens, or if examples show that one of the atoms must be
connected to another lens. The priorities of these lenses change over time, as
each atom can only be involved in one lens. Adding a lens may necessitate
removing another, so not all popped atom lenses will lower the cost of the
generated sequence lens.

@@

We included a number of optimizations in our algorithm. We added ``open'' and
``closed'' annotations to regular expressions, an expansion inference algorithm,
hash-consing and memoization, and compositional synthesis to our algorithm. We
also enable an additional parameter in our specifications that customizes the
termination condition for our procedure.

\paragraph*{Open and Closed Regular Expressions} While GreedySynth acts
relatively efficiently, it suffers from an exponential blowup when converting
regular expressions to DNF form. This can be minimized by not converting
\emph{all} of the regular expression into DNF form, but only converting
\emph{some} of it to DNF form. We add in an additional atom, where an atom can
be an either an iteration of a DNF regular expression or it can be a regular
expression. When converted into a DNF regular expression, if $\Regex$ is a
stochastic regular expression, annotated as closed, then $\ToDNFRegexOf{\Regex}
= \AtomToDNFOf{\Regex}$.

At the start of synthesis, we annotate our regular expressions as closed. Now,
we enhance our rewrites by changing an annotation from closed to open. This
forces changes to \RXSearch and \GreedySynth. \RXSearch now can rewrite closed
regular expressions to open ones. \GreedySynth is provided with the means to map
closed regular expressions to each other. Two closed regular expressions can map
to each other only through the identity lens: two regular expressions can map to
each other by the identity lens, if they are the same closed regular expression
and they parse the same string components in all examples.

\paragraph*{Expansion Inference} The prior optimization helped \GreedySynth, but
it made \RXSearch have a much harder job.  To alleviate this difficulty, we
added in expansion inference.  Expansion inference identifies when a regular
expression does not have a chance to map through any means other than
disconnected.  Expansion inference opens those regular expressions.

@@

\paragraph*{Compositional Synthesis} Compositional synthesis allows us to use
previously defined lenses - whether they are previously defined by users or by
synthesis. We alter \GreedySynth such that two regular expression atoms can be
mapped to each other either if they are the same regular expression and agree
with the examples, or if there is a lens which is a composition of predefined
lenses between them, where the lens agrees with the examples.

>>>>

% begin synthesis
\section{Synthesis}
\label{sec:synthesis}
Algorithm~\ref{alg:synth-sym-lens} presents our synthesis algorithm at a high
level of abstraction. The input regular expressions are first converted into
stochastic regular expressions with \ToStochastic.  
%Once that conversion has
%taken place, the $\RXSearch$ algorithm executes. This algorithm itself is
%defined via five functions: (1) $\PCF{\PQ.Create}$, (2) $\PCF{\PQ.Pop}$, (3)
%$\PCF{Expand}$, (4) $\PCF{\PQ.Push}$, and (5) $\PCF{\Continue}$. First,
This pair of SREs is used to initialize a priority queue ($pq$).
Next, $\PCF{SynthSymLens}$
enters a loop that searches for likely lenses. The loop terminates when the
algorithm believes it is unlikely to find a better lens than the best one it has
found so far (a termination condition defined by $\PCF{\Continue}$). Within each
iteration of the loop, it:
\begin{itemize}
\item pops the next class ($S$, $T$) of lenses to
search off of the priority queue ($\PCF{\PQ.Pop}$),
\item executes $\GreedySynth$ to find a best lens in that class if
one exists ($\LC$),
using the examples $exs$ to filter out potential lenses that do not satisfy
the specification,
\item replaces $best$ with $\LC$, if $\LC$ has lower cost according
to our information theoretic metric, and
\item adds all of the SREs derived from rewriting $S$ and $T$ ($\Expand(S,T)$)
  to the priority queue with $\PCF{\PQ.Push}$ \saz{what determines the priority order?}
\end{itemize}
When the loop terminates, the search returns the globally best lens found
($best$).  Each element of this algorithm will be explained in
further depth in the following subsections.

@@

\subsection{Search Through Types}
The first synthesizer is one that searches through regular expression pairs.
This synthesizer, comprising \Expand, \PQ, and \RXSearch \saz{The name \RXSearch
  does not appear in the figure.  I get the impression that it is the name of
  the while loop, but perhaps we can make that clearer in the algorithm code?
  Maybe by putting a brace around the while loop and giving it a name? Here we
  could say something like ``... and a loop that continues searching until a
  satisfactory lens is found (we call this loop \RXSearch)}.  It proposes regular
expressions that are semantically, but not syntactically, equivalent to the originals. In
particular, this synthesizer performs the following rewrites:
\begin{center}
  \begin{tabular}{rcl}
    $\PRegexStar{\Regex}{\Probability}$
    & \Rewrite
    & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability}\\

@@

\begin{trivlist}
  \centering
\item 
  % \begin{center}
  \begin{tabular}{@{\ }v@{\ }q}
    \LanguageOf{\PRegexStar{\DNFRegex{}}{\Probability}} \ =\  &
                                            \{\String_1\Concat\ldots\Concat\String_n
                                            \SuchThat \forall i,  \String_i\in\LanguageOf{\DNFRegex}\}\\
    \LanguageOf{\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n}}\ =\  & 
    % \\
    % \multicolumn{2}{L}{\ \ \ \ 
    \{\String_0\Concat\StringAlt_1\cdots\StringAlt_n\Concat\String_n \SuchThat \StringAlt_i\in\LanguageOf{\Atom_i}\}
    % }
    \\
    \LanguageOf{\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}}\ =\  &
    % \\
    % \multicolumn{2}{L}{\ \ \ \ 
    \{\String \SuchThat \String \in \LanguageOf{\Sequence_i} \text{\ and $i\in\RangeIncInc{1}{n}$}\}
    % }
  \end{tabular}
  % \end{center}
\end{trivlist}

@@

We have a means of turning stochastic regular expressions into stochastic DNF
regular expressions. The conversion algorithm itself, written
$\ToDNFRegexOf{\Regex}$, is defined below.
\[
  \begin{array}{rcl}
    \ToDNFRegexOf{\String} & = & \DNFOf{(\SequenceOf{\String},1)}\\
    \ToDNFRegexOf{\emptyset} & = & \DNFOf{}\\
    \ToDNFRegexOf{(\PRegexStar{\Regex}{\Probability})} & = & \AtomToDNFOf{\PRegexStar{(\ToDNFRegexOf{\Regex})}{\Probability}}\\
    \ToDNFRegexOf{(\RegexConcat{\Regex_1}{\Regex_2})} & = & \ToDNFRegexOf{\Regex_1} \ConcatDNF \ToDNFRegexOf{\Regex_2}\\
    \ToDNFRegexOf{(\PRegexOr{\Regex_1}{\Regex_2}{\Probability})} & = & \ToDNFRegexOf{\Regex_1} \OrDNF_{\Probability} \ToDNFRegexOf{\Regex_2}\\
  \end{array}
\]
We have proved that this conversion respects languages and probability
distributions.

@@

\paragraph*{\ToStochastic} With stochastic DNF regular expressions and
\ToDNFRegex defined, it is easier to explain the function that converts regular
expressions into stochastic regular expressions, \ToStochastic. If users
specified stochastic regular expressions themselves (perhaps inferred from
example data), this step could be skipped,
though our implementation does not (yet) permit users to provide such
stochastic regular expressions. If $\Regex$ is a stochastic DNF regular
expression generated by \ToStochastic, then $\ToDNFRegexOf{\Regex} =
\DNFOf{(\Sequence_1,\frac{1}{n}) \DNFSep \ldots \DNFSep
  (\Sequence_n,\frac{1}{n})}$ for some sequences $\Sequence_1\ldots\Sequence_n$,
and every stochastic atom generated by \ToStochastic is
$\PRegexStar{\DNFRegex}{.8}$. In particular, $\ToDNFRegex$ generates regular
expressions whose DNF form gives equal probability to all sequence subcomponents
of the DNF regular expression., and gives iterations a .8 chance to continue
iterating. In our experience generating random strings from regular expressions,
these probabilities provide good distributions of strings---stars are iterated
4 times on average, and no individual choice in a series disjunctions is chosen
disproportionately often. As we have found these distributions generate
reasonable strings, we believe they describe the shape of data well.

@@

\saz{The following two paragraphs are pretty similar (was there a bad merge at
  some point?) and very low level.  I also
  could not really follow either of the examples at all.  If we
  are looking for things to cut, I would propose dropping most of the content of
  the next two paragraphs, merging the remainder, and skipping pretty directly to the paragraph
  that starts with ``While DNF lenses...''}
As an example, consider the symmetric DNF lens:
$(\SDNFLensOf{(1,1,\SSQLens_1)\DNFLSep (1,2,\SSQLens_2)}
,\ListOf{2},\ListOf{1;2})$, synchronizing data between $\DNFOf{\Sequence_1}$ and
$\DNFOf{\SequenceAlt_1 \DNFSep \SequenceAlt_2}$. This lens merges both
$\SequenceAlt_1$ and $\SequenceAlt_2$ into $\Sequence_1$, as they both are
involved in a lens with $\Sequence_1$. When performing a \CreateR, $\SSQLens_2$
is used. When performing a \CreateL on a string in $\Sequence_1$, $\SSQLens_1$
is used, and when performing a \CreateL on a string in $\Sequence_2$,
$\SSQLens_2$ is used. More generally, a symmetric DNF Lens \SDNFLens{} consists
of three components. First is the mapping component. The mapping component is a
list of triples $(i,j,\SSQLens)$. In this, $i$ and $j$ refer to sequence $i$ and
sequence $j$ in the left and right DNF regular expressions, respectively. The
third component of the triple, \SSQLens{}, is a sequence lens that maps between
the two sequences specified by $i$ and $j$. The second component is a list of
integers $c$. The integer, $c_i$ at position $i$ states that when performing a
$\CreateR$ on a string matching sequence $\Sequence_i$, it will use the sequence
lens that maps it to sequence $\SequenceAlt_{c_i}$ on the right. The third
component is a list of integers $d$ that follows the same pattern as the second
component, but maps from right to left.

@@

\subsection{\GreedySynth}
\label{subsec:greedy-synth}
\saz{We should call out the structural similarity
  between the DNF grammar and the organization of these algorithms.}
The greedy algorithm is actually two\saz{three! don't forget AtomSynth} greedy algorithms, one that greedily finds
symmetric DNF lenses (\GreedySynth) and one that greedily finds symmetric
sequence lenses (\GreedySeqSynth). We describe this synthesis algorithm in the
context of trying to find a lens between \lstinline{"" | name.name*} and
\lstinline{"" | name}.

\paragraph*{Symmetric DNF Lenses} Algorithm~\ref{alg:greedysynth} presents
\GreedySynth, a means to synthesize symmetric DNF lenses. The inputs to
\GreedySynth are a suite of input-output examples and a pair of stochastic DNF regular expressions.

First, \PCF{CannotMap} checks whether there is no possible lens satisfying the
examples. If so, \GreedySynth returns \None early. Otherwise, \GreedySynth finds
the best lenses (given the examples that match them) between all sequence pairs
$(SQ_i, TQ_j)$ drawn from the left and right DNF regular expressions. (The
function \CartesianMap maps its argument across the cross product of the input
lists).  A priority queue containing these sequence lenses is then initialized
with $\PCF{PQ.Create}$.\saz{Ordered how?} The symmetric lens is then built up
iteratively from these sequence lenses, where the state of the partially constructed lens is
tracked in the lens builder, $lb$.

\GreedySynth loops until there are no more sequence lenses in the priority
queue. Within this loop, a sequence lense is popped from the queue and, if it is
``useful'', then it is included in the final DNF lens. The lens is considered to
be useful when its source (or target) is not already the source (or target) of
an already chosen sequence lens. If the examples require that two sequences have a
lens between them, then any sequence lens between them is considered useful. We
update the priorities of the sequence lenses as the algorithm proceeds: if two
sequence lenses have the same source, the second one to be popped gets a higher
cost than it originally had, to account for the information that needs to be stored
for including that source of non-bijectivity.

In our example, \GreedySynth might first pop the sequence pair \lstinline{""} and
\lstinline{""}, because there is a bijective sequence lens between them. As neither
\lstinline{""} is involved in a sequence lens, this lens is considered useful.
Next, the sequence lens between \lstinline{name.name*} and
\lstinline{name} would be popped: while that lens is not bijective: it is still
better than the alternatives. As all sequences are now involved in sequence lenses,
and there are no examples to make other lenses useful, no further sequence
lenses would be added to the lens builder.

@@

\paragraph*{Symmetric Sequence lenses} Algorithm~\ref{alg:greedyseqsynth}
presents \GreedySeqSynth, which synthesizes symmetric sequence lenses using an
algorithm whose structure is similar to \GreedySynth{}'s.  It calls \PCF{AtomSynth}, a
function that synthesizes atom lenses by iterating a DNF lens between
its subcomponents.

The inputs to \GreedySeqSynth are a suite of input-output
examples and a pair of stochastic atoms. As in \GreedySynth, \GreedySeqSynth
returns \None early if there is no possible lens. Afterward,
\GreedySeqSynth finds the best lenses between each atom pair of the left and
right sequences, and organizes them into a priority queue 
with \PCF{PQ.Create}. The symmetric sequence lens is built up iteratively from
these atom lenses, where the state of the partially built lens is tracked in the sequence lens
builder, $slb$.

\GreedySeqSynth then loops until there are no more atom lenses in the priority
queue. Within this loop, atom lenses are popped from the queue and the
``useful'' ones are kept.  An atom lens is useful if adding it will lower the cost of the
generated sequence lens, or if examples show that one of the atoms must be
connected to another lens.\saz{I'm not sure what ``must be connected to another
  lens means''}  The priorities of these lenses change over time, because
each atom can only be involved in one lens. Adding a lens may necessitate
removing another, so not all popped atom lenses will lower the cost of the
generated sequence lens.\saz{I don't understand the last two sentences.}

@@

Our implementation includes a number of optimizations to this algorithm.  We added ``open'' and
``closed'' annotations to regular expressions \saz{what do these mean? I guess
  ``open'' means ``subject to  DNF conversion'' and ``closed'' means not?}, an expansion inference algorithm,
hash-consing and memoization, along with compositional synthesis. The implementation
also allows an additional parameter in specifications that customizes the
termination condition for the synthesis procedure.

\paragraph*{Open and Closed Regular Expressions} While \GreedySynth acts
relatively efficiently, it suffers from an exponential blowup when converting
regular expressions to DNF form. This can be minimized by not converting
\emph{all} of the regular expression into DNF form, but only converting
\emph{some} of it to DNF form, and doing so incrementally.  This change is easy
to make by adding non-DNF stochastic regular expressions as a new type of
``closed'' atom.  When converted into a DNF regular expression, if $\Regex$ is
such a closed stochastic regular expression, then
$\ToDNFRegexOf{\Regex} = \AtomToDNFOf{\Regex}$.

At the start of synthesis, all regular expressions are closed. \RXSearch is
modified so that it is now permitted to rewrite closed regular expressions to
open ones (thereby triggering DNF normalization). \GreedySynth is allowed to to
map closed regular expressions to each other, but only using the identity
lens---they must parse the same string components in all examples.

\paragraph*{Expansion Inference} The prior optimization helped \GreedySynth, but
it made \RXSearch have the harder job of determining when to open closed
expressions.  To alleviate this difficulty, we added in expansion inference,
which identifies when a regular expression cannot possibly map by any means
other than the disconnect lens.  Expansion inference opens those regular
expressions to allow for further exploration.

@@

\paragraph*{Compositional Synthesis} Compositional synthesis allows us to use
previously defined lenses---whether they are previously defined by users or by
synthesis. We alter \GreedySynth such that two regular expression atoms can be
mapped to each other either if they are the same regular expression and agree
with the examples, or if there is a lens which is a composition of predefined
lenses between them, where the lens agrees with the examples.
